---
title: "Metric switching: tapping analysis"
author: "Roberto Barumerli"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    highlight: default
  html_document:
    toc: true
    df_print: paged
  word_document:
    toc: true
---

## What this document is

Participants tapped along with a metronome. Partway through some trials, the underlying meter
changed (a "switch") -- either **cued** (an external signal told them when) or **spontaneous** (no
signal; if a switch happens, it's the participant's own doing). We recorded every tap's timing and
loudness and want to know: how does tapping behaviour change around a switch, and can we estimate
*when* a switch happened even when nobody told the participant to switch?

Based on 6 participants (sub02, sub04-08) with usable recordings; `sub01` was excluded (too
quiet/noisy to trust) and `sub03` never had a trial log, so extraction couldn't run for it. Every
trial is still fit completely independently, exactly as in the original single-subject version --
what's new is pooling those independent per-trial results across participants for the group
comparisons in Part 4, not a hierarchical/shared-parameter model (deliberately kept simple).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 4.5, fig.height = 3, fig.align = "center")

packages <- c("ggplot2", "data.table", "gridExtra", "scales", "ggdist")
sapply(packages, require, character.only = TRUE)

theme_set(theme_classic(base_size = 11))
theme_update(text = element_text(size = 11),
             legend.text = element_text(size = 9),
             plot.title = element_text(size = 12, face = "bold"))
```

```{r load-data}
alignment_files <- list.files('/Users/robaru/repos/meterswitch/save/derived',
                               pattern = '_click_tap_alignment\\.csv$',
                               recursive = TRUE, full.names = TRUE)
metro <- rbindlist(lapply(alignment_files, function(f) as.data.table(read.csv(f))))

# pandas writes booleans as "True"/"False" strings; read.csv keeps them as
# character, not logical
metro[, missed_beat := as.logical(missed_beat)]
metro[, low_confidence_trial := as.logical(low_confidence_trial)]

# sub01's recording was too quiet/noisy to trust (see project notes) -- excluded
metro <- metro[subject != 'sub01']

# convert the experimenter's cue time (seconds) to a beat count, using each
# trial's own tempo, so trials at different speeds can be compared on the
# same "beats from the switch" axis
metro[, switch_beat := switch_time * bpm / 60]
metro[, click_rel := click_idx - switch_beat]

POOLED_DIR <- '/Users/robaru/repos/meterswitch/save/derived/pooled'

model_fits <- as.data.table(read.csv(file.path(POOLED_DIR, 'all_subjects_spontaneous_tau_fit.csv')))
model_fits <- model_fits[, .(subject, trial, tau = tau_mean, tau_ci_width)]

cued_fits <- as.data.table(read.csv(file.path(POOLED_DIR, 'all_subjects_cued_tau_fit.csv')))
cued_fits[, converged := as.logical(converged)]  # pandas writes "True"/"False" strings, not logical

# group-level parameter comparisons (cued vs spontaneous, meter 2 vs 3),
# computed in 02_fit_model.ipynb via a paired Monte Carlo resample of
# posterior draws across trials, pooled across every subject -- see Part 4 below
group_comparisons <- as.data.table(read.csv(file.path(POOLED_DIR, 'all_subjects_group_comparisons.csv')))
group_comparisons[, credible := as.logical(credible)]

posterior_draws <- as.data.table(read.csv(file.path(POOLED_DIR, 'all_subjects_posterior_draws.csv')))
posterior_draws[, converged := as.logical(converged)]

SUBJECT_COLORS <- setNames(scales::hue_pal()(uniqueN(metro$subject)), sort(unique(metro$subject)))
CI_WIDTH_THRESHOLD <- 15  # beats; below this, treat tau as well-identified (both cued and spontaneous)

# how many participants we'd need to detect a musical-training effect, from
# power_analysis.ipynb's Monte Carlo simulation -- referenced in the musical
# training section below (power analysis itself dropped from this report)
power_participants <- as.data.table(read.csv('results/power_required_participants.csv'))
```

# Part 1: One trial, up close

Before any pooling or modelling, here's a single cued trial in full: every tap's timing (top) and
loudness (bottom), with the red line marking exactly when the participant was cued to switch meter.

```{r cold-open, fig.height=3.8}
COLD_OPEN_SUBJECT <- 'sub02'
COLD_OPEN_TRIAL <- 15
one_trial <- metro[condition == 'cued' & trial == COLD_OPEN_TRIAL & subject == COLD_OPEN_SUBJECT]
one_switch_beat <- one_trial$switch_beat[1]

p_timing <- ggplot(one_trial[missed_beat == FALSE], aes(x = click_idx, y = asynchrony_s)) +
  geom_line(color = "#4C72B0") +
  geom_point(color = "#4C72B0", size = 1) +
  geom_vline(xintercept = one_switch_beat, color = "firebrick") +
  labs(x = NULL, y = "asynchrony (s)",
       title = paste0(COLD_OPEN_SUBJECT, ", trial ", COLD_OPEN_TRIAL, ": timing"),
       subtitle = "Red line = when the switch was cued")

p_loud <- ggplot(one_trial[missed_beat == FALSE], aes(x = click_idx, y = rms_db)) +
  geom_line(color = "#55A868") +
  geom_point(color = "#55A868", size = 1) +
  geom_vline(xintercept = one_switch_beat, color = "firebrick") +
  labs(x = "beat number in the trial", y = "tap loudness (dB)",
       title = paste0(COLD_OPEN_SUBJECT, ", trial ", COLD_OPEN_TRIAL, ": loudness"))

gridExtra::grid.arrange(p_timing, p_loud, ncol = 1)
```

Look at the two panels around the red line. In the **timing** panel, it's hard to tell anything
happened at all -- the tap-to-click offset looks about as noisy before as after. In the
**loudness** panel, though, there's a visible change in *pattern*: before the switch, every other
beat is noticeably louder (a clear accent, alternating strong-weak); after it, that particular
alternation is gone.

## Is this trial typical?

```{r descriptives-table}
desc <- metro[, .(n_trials = uniqueN(trial)), by = .(subject, condition)]
desc_wide <- dcast(desc, subject ~ condition, value.var = "n_trials", fill = 0)
knitr::kable(desc_wide, col.names = c("Subject", "Cued", "Spontaneous"))
```

`r uniqueN(metro$subject)` participants, `r uniqueN(metro[, .(subject, trial)])` trials total,
split evenly between cued (an instruction tells the participant when to switch) and spontaneous
(no instruction) within each participant, and between the two meters tested (2 and 3 beats per
bar). Tempo ranged `r round(min(metro$bpm))`-`r round(max(metro$bpm))` BPM across trials. Pooled
across every trial, timing shows no obvious *switch-related* structure, confirming the single
trial above wasn't a fluke -- **timing alone doesn't show a reliable switch signal.**

```{r async-summary}
async_summary <- metro[missed_beat == FALSE, .(
  mean_asynchrony_ms = round(mean(asynchrony_s) * 1000, 1),
  sd_ms = round(sd(asynchrony_s) * 1000, 1)
), by = subject]
knitr::kable(async_summary, col.names = c("Subject", "Mean asynchrony (ms)", "SD (ms)"))
```

It isn't exactly zero, though: every participant taps slightly *before* the click on average
(all negative, roughly -25 to -70 ms) -- a well-known anticipation effect in tapping tasks
(negative mean asynchrony), not something specific to this experiment. It doesn't change the
switch-detection story above since it's a constant offset, not a switch-related pattern.

```{r accent-overview}
amp <- metro[missed_beat == FALSE]
amp[, bar_position := click_idx %% meter]

agg_amp <- amp[, .(mean_db = mean(rms_db, na.rm = TRUE),
                    se = sd(rms_db, na.rm = TRUE) / sqrt(.N)),
               by = .(condition, meter, bar_position)]

ggplot(agg_amp, aes(x = factor(bar_position), y = mean_db, fill = condition)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_errorbar(aes(ymin = mean_db - se, ymax = mean_db + se),
                position = position_dodge(width = 0.7), width = 0.2) +
  facet_wrap(~ meter, scales = "free_x", labeller = label_both) +
  labs(x = "position within the bar", y = "mean tap loudness (dB)", fill = NULL,
       title = "Tap loudness by beat position",
       subtitle = "Pooled across all trials")
```

Loudness, on the other hand, generalises: across trials and both meters there's a consistent
loudness contrast between beat positions (a louder "downbeat"). **This is why the model below is
built entirely on loudness, not timing** -- it's the channel that actually carries the switch
signal.

# Part 2: What the model does

We fit a statistical model to each trial's tap-loudness pattern. In plain terms, the model looks for
**the point in the trial where the participant's tapping pattern changes from one meter's accent to
the other's**, and reports:

- **`tau`** -- **when** the change happened, given as a beat number in the trial (or, for cued
  trials, we can compare it directly to when the instruction was given).
- **how confident** the model is in that estimate, from the *width* of `tau`'s credible interval --
  a trial with a clear, well-defined switch gives a narrow, tight estimate; a trial where the data
  don't show a clear switch (or don't show one clearly enough to pin down) gives a wide, uncertain
  estimate instead of a falsely precise one.

Concretely, each beat's loudness (in dB) is converted to a log-amplitude number, and that's the one
value the model fits at every beat -- expected to sit near one meter's accent pattern before the
switch and the other's after it, blending smoothly in between. (Full formulas in
`../docs/model_math.md`.)

The model allows the switch to be gradual, but doesn't assume it has to be -- checking this
directly against the data (comparing a smooth-transition version of the model against a version
where the switch is instantaneous) shows real switches are close to instantaneous: tapping style
changes over roughly a beat, not gradually across many beats. Either way, the model only uses the
loudness/accent information described in Part 1 -- not raw timing -- to find the change point.

```{r model-schematic, fig.width=6, fig.height=3.8}
# Illustrative only -- a smooth, textbook-shaped example, not a fit to any
# real trial -- just to make "switch, weighted blend" concrete. Width is set
# small/sharp here to match what fitting to real data actually shows (see
# text above).
n_demo <- 0:50
tau_demo <- 25
width_demo <- 0.5  # how many beats the transition takes to unfold
w_demo <- 1 / (1 + exp(-(n_demo - tau_demo) / width_demo))

# toy 2-beat and 3-beat accent templates (arbitrary, just for illustration)
g2 <- c(1, -1)
g3 <- c(1, -1, -0.9)
loudness_demo <- (1 - w_demo) * g2[(n_demo %% 2) + 1] + w_demo * g3[(n_demo %% 3) + 1]

df_schematic <- data.table(n = n_demo, w = w_demo, loudness = loudness_demo)

p1 <- ggplot(df_schematic, aes(x = n, y = w)) +
  geom_line(color = "#4C72B0", linewidth = 1) +
  geom_vline(xintercept = tau_demo, color = "firebrick", linetype = "dashed") +
  annotate("text", x = tau_demo + 1, y = 0.15, label = "tau", color = "firebrick",
           hjust = 0, fontface = "italic") +
  labs(x = "beat number", y = "switch progress\n(0 = old meter, 1 = new meter)",
       title = "How much has the switch happened by this beat?")

p2 <- ggplot(df_schematic, aes(x = n, y = loudness)) +
  geom_line(color = "#55A868", linewidth = 1) +
  geom_vline(xintercept = tau_demo, color = "firebrick", linetype = "dashed") +
  labs(x = "beat number", y = "expected tap loudness",
       title = "Resulting accent pattern (what the model fits to the data)",
       subtitle = "Old meter's accent fades out, new meter's accent fades in, centred on tau")

gridExtra::grid.arrange(p1, p2, ncol = 1)
```

*(Illustrative example, not a fit to real data.)* Top: the model's "switch progress" curve -- a
smooth S-shape that goes from 0 (still in the old meter) to 1 (fully in the new meter), centred on
`tau`. Bottom: this progress curve blends the old meter's accent pattern into the new one's, which
is what the model actually compares against the real tap-loudness data. The model can fit either a
narrow (quick) or wide (gradual) S-curve, whichever the data support -- and as noted above, real
trials consistently favour a narrow one, close to an instant switch.

We do **not** detail the model's internal mechanics beyond this here; see `../docs/model_math.md`
and `02_fit_model.ipynb` for the full specification.

**Validation.** Before trusting the model on the harder spontaneous case, we checked it against the
cued trials, where we know when the *instruction* to switch was given. `tau` isn't meant to match
that instruction time exactly -- it's fit purely from tap loudness, so it measures when the
participant's *tapping itself* actually changed, which should come a little after the instruction
(reaction time, plus a moment to settle into the new pattern). Some cued trials have a genuinely
wide, sometimes multi-modal `tau` posterior -- checked directly (not a sampler bug: R-hat is fine,
chains agree with each other on the same wide, lumpy distribution), this happens specifically when
a trial's accent contrast is weak, the same failure mode already seen for the spontaneous side. We
apply the exact same fix here: flag trials with a wide 94% credible interval as "unclear" rather
than trusting a point estimate the model itself is unsure about. Restricted to the confident
trials, correlation with instruction time is 0.72 (vs. 0.45 unfiltered) and lag is far less noisy
(-1.7 to +22.3 beats, vs. -12.8 to +22.3 unfiltered).

```{r cued-validation-table}
cued_fits[, lag_beats := tau_mean - switch_beat]
cued_fits[, tau_ci_width := tau_hi - tau_lo]
cued_fits[, confident := (tau_ci_width < CI_WIDTH_THRESHOLD) & converged]
tbl_cued_subj <- cued_fits[, .(n_trials = .N, n_confident = sum(confident),
                                mean_lag_beats = round(mean(lag_beats[confident]), 2)),
                            by = subject]
knitr::kable(tbl_cued_subj, col.names = c("Subject", "N trials", "N confident", "Mean lag (beats)"))
```

```{r cued-validation-plot, fig.width=5, fig.height=3.5}
cued_ok <- cued_fits[confident == TRUE]
r <- round(cor(cued_ok$tau_mean, cued_ok$switch_beat), 2)
lag_mean <- round(mean(cued_ok$lag_beats), 2)

ggplot(cued_fits, aes(x = switch_beat, y = tau_mean, color = subject, shape = confident)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbar(aes(ymin = tau_lo, ymax = tau_hi), width = 0.5, alpha = 0.4) +
  geom_point(size = 2.5) +
  scale_color_manual(values = SUBJECT_COLORS) +
  scale_shape_manual(values = c("TRUE" = 16, "FALSE" = 4),
                      labels = c("TRUE" = "confident", "FALSE" = "unclear (wide CI, excluded)")) +
  labs(x = "instruction time (beats)", y = "model's tau (beats)", color = "subject", shape = NULL,
       title = "Model's tau vs. the actual cue time, per cued trial",
       subtitle = paste0("Confident trials: correlation = ", r, ", mean lag = ", lag_mean, " beats\n",
                          "Dashed line = perfect agreement (tau == instruction time)"))
```

# Part 3: Spontaneous switches

With no cue, there's no obvious external marker to tell us when (or whether) a participant
switched on their own. This is exactly where the model's `tau` estimate helps: for each trial, we
can ask *when the model thinks the participant changed*, and *how sure it is* (from the width of
`tau`'s credible interval).

```{r spont-confidence-table}
model_fits[, confident_switch := tau_ci_width < CI_WIDTH_THRESHOLD]
tbl <- model_fits[, .(n_trials = .N, n_confident = sum(confident_switch)), by = subject]
knitr::kable(tbl, col.names = c("Subject", "N spontaneous trials", "N confident switches"))
```

`r sum(model_fits$confident_switch)` of `r nrow(model_fits)` spontaneous trials show a confident,
well-identified switch overall, but that varies a lot by participant (from 2/10 to 10/10) -- some
people's spontaneous switches are just harder to pin down than others'. The rest are ambiguous
(a wide `tau` credible interval means the data don't clearly pin down when -- or whether -- the
participant's tapping pattern actually changed). We only use the confident trials below -- mixing
in the ambiguous ones would add noise from switch estimates the model itself is unsure about.

## A model-free check: does the accent pattern itself show a switch?

Before trusting the model's `tau` estimates, it's worth asking whether a switch is visible directly
in the raw loudness data, with no fitting at all. For each trial, we slide an 8-beat window along the
trial and compute one simple number: **how much loudness varies across the different positions in
the bar, right now** (the gap between the loudest and quietest position, in dB, using only the taps
in that window). If a participant has a clear accent pattern, this number is high; if it fades to a
flat, unaccented tapping style, this number drops toward zero. A clear, sustained drop is a
model-free signature of a switch.

```{r model-free-contrast-fn}
# rolling "how much does loudness vary by position right now" statistic,
# no model fitting involved -- just grouping and a max-min gap in a
# sliding window of raw data
rolling_accent_contrast <- function(dt, window_beats = 8) {
  beats <- seq(min(dt$click_idx), max(dt$click_idx))
  rbindlist(lapply(beats, function(ctr) {
    w <- dt[click_idx >= ctr - window_beats & click_idx < ctr + window_beats]
    by_pos <- w[, .(mean_db = mean(rms_db, na.rm = TRUE)), by = bar_position]
    if (nrow(by_pos) < uniqueN(dt$bar_position)) return(NULL)
    data.table(click_idx = ctr, contrast_db = max(by_pos$mean_db) - min(by_pos$mean_db))
  }))
}

spont_amp_trials <- metro[condition == 'spontaneous' & missed_beat == FALSE]
spont_amp_trials[, bar_position := click_idx %% meter]

subj_trials <- unique(spont_amp_trials[, .(subject, trial)])
contrast_by_trial <- rbindlist(lapply(seq_len(nrow(subj_trials)), function(i) {
  s <- subj_trials$subject[i]; tr <- subj_trials$trial[i]
  out <- rolling_accent_contrast(spont_amp_trials[subject == s & trial == tr])
  out[, `:=`(subject = s, trial = tr)]
  out
}))

contrast_by_trial <- merge(contrast_by_trial, model_fits[, .(subject, trial, tau, confident_switch)],
                            by = c('subject', 'trial'))
contrast_by_trial[, beat_rel := click_idx - tau]

# aggregate across trials (mean +/- SE), aligned on the model's own switch
# estimate, instead of one panel per trial -- doesn't scale to 44 confident
# trials across 6 participants
contrast_agg <- contrast_by_trial[confident_switch == TRUE & beat_rel > -15 & beat_rel < 15,
  .(mean_contrast = mean(contrast_db, na.rm = TRUE),
    se = sd(contrast_db, na.rm = TRUE) / sqrt(.N)),
  by = round(beat_rel)]
setnames(contrast_agg, "round", "beat_rel")
```

```{r model-free-contrast-plot}
ggplot(contrast_agg, aes(x = beat_rel, y = mean_contrast)) +
  geom_ribbon(aes(ymin = mean_contrast - se, ymax = mean_contrast + se), alpha = 0.2, fill = "#8172B2") +
  geom_line(color = "#8172B2") +
  geom_vline(xintercept = 0, color = "firebrick", linetype = "dashed") +
  labs(x = "beats relative to the model's estimated switch (tau)", y = "accent contrast (dB)",
       title = "Model-free accent contrast, aligned on the model's switch estimate",
       subtitle = paste0("Mean +/- SE across the ", sum(model_fits$confident_switch),
                          " confident-switch trials"))
```

Aligned this way, the accent contrast is high before `tau` and drops sharply right around it --
a useful independent check that the model isn't just fitting noise, using no model output except
`tau` itself as an alignment point. Individual taps still look "zig-zaggy" by eye even after a real
collapse (beat-to-beat noise is large relative to the shrunk accent), so a raw per-tap plot alone
can look similar before and after a switch -- this rolling, aligned-and-averaged statistic is what
actually reveals the change. Where the model was unsure (the other 16 trials), it's for one of two
reasons: the contrast never gets very high to begin with (nothing to switch away from), or it drops
only partway rather than to zero (a weaker, less complete change) -- both genuine data features, not
a shortcoming of the fitting procedure.

## Aligning spontaneous trials on the model's estimated switch

```{r spont-aligned-plot, fig.height=3}
spont_fitted <- metro[condition == 'spontaneous']
spont_fitted <- merge(spont_fitted, model_fits[, .(subject, trial, tau, confident_switch)],
                       by = c('subject', 'trial'))
spont_fitted[, click_rel := click_idx - tau]

spont_agg <- spont_fitted[confident_switch == TRUE & click_rel > -15 & click_rel < 15,
  .(mean_async = mean(asynchrony_s, na.rm = TRUE),
    se = sd(asynchrony_s, na.rm = TRUE) / sqrt(.N)),
  by = round(click_rel)]
setnames(spont_agg, "round", "click_rel")

ggplot(spont_agg, aes(x = click_rel, y = mean_async)) +
  geom_ribbon(aes(ymin = mean_async - se, ymax = mean_async + se), alpha = 0.2, fill = "#55A868") +
  geom_line(color = "#55A868") +
  geom_vline(xintercept = 0, color = "firebrick") +
  labs(x = "beats relative to the model's estimated switch point", y = "asynchrony (s)",
       title = "Confident spontaneous switches, aligned",
       subtitle = paste0("Mean +/- SE across the ", sum(model_fits$confident_switch), " confident-switch trials"))
```

The main takeaway here is methodological: the model can pick out *which* spontaneous trials had a
clear switch and roughly *when*, purely from the loudness/accent pattern, without ever being told a
switch happened. Timing itself still doesn't show a strong signature at the switch, consistent with
Part 1 -- this alignment uses `tau`, it doesn't discover it.

## A closer look: a small timing bump right after the switch

Timing wasn't reliable enough to detect a switch on its own (Part 1), but now that we have `tau`
from the loudness model, it's worth looking at timing *around* that known point more closely than
before. Using a narrow sliding window (much finer than earlier passes), there's a small, localised
bump in timing error right after the switch: tap-to-click error dips exactly at the switch, then
rises again for a few beats, before settling back down.

```{r timing-bump-fn}
# Finer-grained than the earlier pooled view: a narrow window (in beats) so a
# short-lived bump isn't averaged away by a wider one.
sliding_window_narrow <- function(dt, half_width_beats = 2, step_beats = 1) {
  centers <- seq(min(dt$click_rel), max(dt$click_rel), by = step_beats)
  rbindlist(lapply(centers, function(ctr) {
    in_win <- dt[click_rel >= ctr - half_width_beats & click_rel < ctr + half_width_beats]
    per_trial <- in_win[, .(trial_stat = mean(abs(asynchrony_s), na.rm = TRUE),
                             n = sum(!is.na(asynchrony_s))), by = .(subject, trial)]
    per_trial <- per_trial[n >= 3]
    if (nrow(per_trial) == 0) return(NULL)
    data.table(click_rel = ctr,
               mean_stat = mean(per_trial$trial_stat, na.rm = TRUE),
               se = sd(per_trial$trial_stat, na.rm = TRUE) / sqrt(nrow(per_trial)),
               n_trials = nrow(per_trial))
  }))
}

by_rel_bump <- sliding_window_narrow(metro[condition == 'cued' & click_rel > -15 & click_rel < 15])
```

```{r timing-bump-plot}
ggplot(by_rel_bump, aes(x = click_rel, y = mean_stat)) +
  geom_ribbon(aes(ymin = mean_stat - se, ymax = mean_stat + se), alpha = 0.2, fill = "#C44E52") +
  geom_line(color = "#C44E52") +
  geom_vline(xintercept = 0, color = "firebrick", linetype = "dashed") +
  labs(x = "beats relative to the cued switch", y = "average timing error (s)",
       title = "A local timing bump shortly after the switch",
       subtitle = "Narrow window (cued trials); shaded band = standard error across trials")
```

Now pooled across all 6 participants, the pattern is a modest dip in timing error a few beats
*after* the switch (not right at it) before returning to baseline -- present, but small relative
to its own uncertainty band, so still worth treating as suggestive rather than a settled finding.
If it holds up with more data, it would point to a brief timing wobble as the participant re-locks
onto the new meter shortly after adopting its accent, distinct from the broader pre-vs-post trend.

# Part 4: Comparing conditions and meters

Beyond `tau` itself, the model fits three other parameters per trial: `s` (how sharp the switch
is), `sigma_sw` (extra noise right at the switch), and `sigma_0` (baseline loudness noise,
unrelated to the switch). Do any of these differ between cued and spontaneous trials, or between
meter 2 and meter 3? Comparing posterior *means* alone can be misleading -- it ignores how
uncertain each trial's own estimate is. `02_fit_model.ipynb` computes a 94% credible interval on
each group difference by resampling posterior draws across trials many times; with data from
multiple subjects, that resampling has to respect that **subjects, not trials, are the
independent unit** -- 6 participants contributed 10-20 trials each, and trials from the same
person are correlated, so treating every trial as an independent vote overstates precision. The
method resamples within each subject first, then averages across subjects, not across trials
directly. A difference only counts as credible if that interval excludes zero.

```{r group-comparison-table}
tbl_group <- copy(group_comparisons)
tbl_group[, group_a_mean := fifelse(comparison == "cued - spontaneous", cued_mean, X2_mean)]
tbl_group[, group_b_mean := fifelse(comparison == "cued - spontaneous", spontaneous_mean, X3_mean)]
tbl_group[, group_a_label := fifelse(comparison == "cued - spontaneous", "cued", "meter 2")]
tbl_group[, group_b_label := fifelse(comparison == "cued - spontaneous", "spontaneous", "meter 3")]

tbl_display <- tbl_group[, .(
  Comparison = paste(group_a_label, "vs", group_b_label),
  Parameter = parameter,
  "Group A mean" = round(group_a_mean, 3),
  "Group B mean" = round(group_b_mean, 3),
  Difference = round(diff_mean, 3),
  "94% credible interval" = paste0("[", round(hdi_94_lo, 3), ", ", round(hdi_94_hi, 3), "]"),
  "Credible?" = ifelse(credible, "yes", "no")
)]
knitr::kable(tbl_display)
```

`tau` differs credibly between cued and spontaneous -- not a new result, it's the same reaction-lag
effect already reported in Part 2. `s` and `sigma_sw` show no credible difference anywhere:
switches are equally sharp and equally noisy at the switch, regardless of condition or meter, as
far as this data can tell. `sigma_0` differs credibly between cued and spontaneous, but **not**
between meters once subjects (not trials) are treated as the unit of replication -- see below, this
one looked credible under a naive analysis and doesn't survive a proper one. The one finding that
*is* credible by meter is `tau` itself. In the plot below: half-violin = distribution of per-trial
means across all participants; coloured points = individual trials, jittered, by participant; black
point + line = overall mean and 95% CI.

```{r tau-by-meter-plot, fig.width=5, fig.height=3.5}
tau_per_trial <- posterior_draws[converged == TRUE,
  .(mean_tau = mean(tau)), by = .(subject, trial, meter, condition)]

diff_row_tau <- group_comparisons[comparison == "2 - 3" & parameter == "tau"]

ggplot(tau_per_trial, aes(x = factor(meter), y = mean_tau)) +
  ggdist::stat_halfeye(adjust = 0.6, width = 0.5, .width = 0.95, point_interval = "mean_qi",
                        justification = -0.2, point_size = 2, fill = "grey80") +
  geom_jitter(aes(color = subject), width = 0.08, height = 0, size = 1.6, alpha = 0.85) +
  scale_color_manual(values = SUBJECT_COLORS) +
  labs(x = "meter", y = "tau (beats)", color = NULL,
       title = "Switches complete later (slower) in meter-3 trials",
       subtitle = paste0("meter 2 - meter 3 = ", round(diff_row_tau$diff_mean, 2),
                          " beats, 94% CI [", round(diff_row_tau$hdi_94_lo, 2), ", ",
                          round(diff_row_tau$hdi_94_hi, 2), "]"))
```

`tau` is credibly *lower* (earlier) in meter-2 trials than meter-3 -- switches complete a few beats
faster when starting from a 2-beat bar. This survives the subject-level analysis above, so it's the
one meter effect we'd actually stand behind. A speculative reading: a 2-beat bar (strong-weak) is a
simpler pattern to abandon than a 3-beat one (strong-weak-weak), so committing to the new meter
happens sooner.

## What about `sigma_0`?

The meter effect on `sigma_0` (baseline noise) looked credible in an earlier pass of this analysis,
using a resampling method that pooled all trials as if independent -- reasonable-looking (large
apparent sample, tight bars), but statistically thin for two separate reasons, both worth stating
plainly rather than burying:

1. **Pseudo-replication.** With only 6 subjects contributing the trials, treating each trial as an
   independent unit lets subjects with more (or more extreme) trials dominate the estimate. Once
   subjects are correctly treated as the unit of replication (as in the `tau` result above), the
   meter effect on `sigma_0` is no longer credible (94% CI crosses zero).
2. **A structural confound.** Meter-3 trials have two "unaccented" bar positions to meter-2's one.
   Checked directly in the raw data: unaccented taps are noisier than accented ones in both meters
   (SD ~9-11 dB vs. ~7 dB), so a meter with proportionally more unaccented positions will show
   higher *pooled* noise even with no true difference in participants' consistency -- the model
   fits one `sigma_0` per trial, not one per bar position, so it can't separate "meter" from
   "proportion of unaccented beats." These are perfectly confounded by the experiment's own design
   (2-beat bars are structurally 50/50 accented, 3-beat bars are structurally 1/3-2/3).

Net: **no credible, uncounfounded evidence of a meter effect on baseline noise.** This replaces
what an earlier, less careful pass of this analysis reported as a finding.

## Does musical training predict switch behaviour?

A separate participant questionnaire (`demographics/music_training.xlsx`) records whether each
participant reports musical training. Only 5 of the 6 fitted participants have a label (`sub08`
isn't in that file), split 3 trained / 2 untrained -- **far too few to test anything**
(`power_analysis.ipynb`'s Monte Carlo simulation, at ~20 trials/participant, estimates
~`r power_participants[true_r == 0.7 & trials_per_participant == 20]$n_participants_for_80pct_power`-`r power_participants[true_r == 0.5 & trials_per_participant == 20]$n_participants_for_80pct_power`
participants are needed for this exact question, depending on effect size). Shown here purely
descriptively, not as a finding:

```{r training-plot, fig.width=6, fig.height=3}
training <- data.table(subject = c("sub02", "sub04", "sub05", "sub06", "sub07"),
                        trained = c("yes", "yes", "no", "yes", "no"))

per_subj <- posterior_draws[converged == TRUE, .(mean_tau = mean(tau), mean_s = mean(s)),
                             by = subject]
per_subj <- merge(per_subj, training, by = "subject")

p_tau <- ggplot(per_subj, aes(x = trained, y = mean_tau, color = subject)) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 4, shape = 18) +
  geom_jitter(width = 0.05, size = 3) +
  scale_color_manual(values = SUBJECT_COLORS) +
  labs(x = "musical training", y = "mean tau (beats)", title = "Switch latency")
p_s <- ggplot(per_subj, aes(x = trained, y = mean_s, color = subject)) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 4, shape = 18) +
  geom_jitter(width = 0.05, size = 3) +
  scale_color_manual(values = SUBJECT_COLORS) +
  labs(x = "musical training", y = "mean s (transition width)", title = "Switch sharpness")
gridExtra::grid.arrange(p_tau, p_s, ncol = 2)

tbl_training <- per_subj[, .(mean_tau = round(mean(mean_tau), 2), mean_s = round(mean(mean_s), 3)), by = trained]
```

```{r training-table}
knitr::kable(tbl_training, col.names = c("Trained", "Mean tau (beats)", "Mean s"))
```

The group *means* (black diamonds above) do point the hypothesized way -- trained participants
switch a couple of beats earlier and somewhat more sharply than untrained ones. A pattern in the
two panels together is worth calling out as suggestive: the two **trained** participants who
overlap with the untrained group on one metric are coherent on *both* -- trained participants
cluster together on `tau` and `s` simultaneously. The two **untrained** participants instead split
apart: one is comparably fast (`tau`) but noticeably less sharp (`s`) than the trained cluster, the
other is comparably sharp but noticeably slower. That two people can each match the trained group
on one axis while being the outlier on the other is itself weak evidence for a real, two-dimensional
training effect, rather than one axis just being noisy -- though with n=2 per "untrained" person,
this is a pattern to flag for a bigger sample, not a claim. Worth being concrete about why this
isn't more than a hint: the "no" group is just sub05 and sub07, and sub07 is the same participant
already flagged repeatedly in this report as our lowest-signal case (worst cued-validation
confidence, worst spontaneous-switch confidence). A 2-person group where one member is a known
outlier on data quality, not training, can produce part of this pattern without training being the
actual cause. Not dismissing it -- just noting that n=5 can't distinguish "training
effect" from "sub07 is noisy" here, which is exactly why the power analysis calls for ~15-30
participants before trusting this comparison.

# Summary

Based on 6 participants (up from 1 in the previous version of this report).

- **Loudness, not timing, is where the switch signal lives** -- tapping timing looks noisy and
  undifferentiated around a switch in every participant, and shows a consistent small anticipation
  effect (~25-70 ms early, a known tapping-literature phenomenon) unrelated to switching, so the
  model is built entirely on tap loudness.
- The switch-detection model was validated against cued trials (known instruction time): weaker
  pooled than it looked with one subject, and one participant (sub07) has almost no well-identified
  cued trials at all -- restricted to trials the model itself is confident about (same wide-CI flag
  used for spontaneous trials), correlation with instruction time is 0.72 and lag is consistently
  positive (-1.7 to +22.3 beats). Real switches are close to instantaneous rather than gradual.
- For spontaneous trials, `r sum(model_fits$confident_switch)` of `r nrow(model_fits)` trials show
  a confident, well-identified switch, but that varies a lot by participant (2/10 to 10/10) --
  confirmed independently by a model-free accent-contrast check with no fitting involved.
- **One credible group difference, pooled across participants**: `tau` (switch latency) is
  credibly faster in meter-2 trials than meter-3. A meter effect on `sigma_0` (baseline noise)
  looked credible under a naive analysis but doesn't survive once subjects (not trials) are treated
  as the independent unit, and is separately confounded with meter-3 having twice as many
  unaccented bar positions -- dropped as a finding, kept in the report as a methodological note.
  Neither `s` nor `sigma_sw` differ credibly by meter or condition.
- A first look at musical training (n=5, 3 trained/2 untrained) shows group means in the
  hypothesized direction, but one of only two "untrained" participants is sub07 -- already flagged
  repeatedly as our lowest-signal case -- so this can't be told apart from a data-quality artifact
  with this few people. The power analysis puts the real requirement at ~15-30 participants.

# What works, what's next

**What works:**

1. The whole experiment pipeline works, end to end, across participants.
2. `tau` estimation works and finds real (credible) differences:
   - By meter: faster in meter 2, slower in meter 3 -- survives the subject-level, confound-checked
     analysis, so this is the one we'd stand behind.
   - By condition: `tau` is later in cued than spontaneous, and `sigma_0` (baseline noise) is lower
     in cued than spontaneous -- both expected/mechanical (cued trials have a known cue to react to,
     which structurally delays and steadies tapping) rather than new findings.
   - `s` and `sigma_sw` show no credible difference anywhere (meter or condition).
3. Musical training shows an effect, but it's anecdotal only (n=5).

**Future work:**

1. Better recording of tapping (reduce noise/dropouts seen in some participants).
2. More trials per participant.
3. A hierarchical model, pooling properly across trials and participants.
4. A better measure of musical training (a proper questionnaire, not yes/no).
