---
title: "Metric switching: tapping analysis"
author: "Roberto Barumerli"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    highlight: default
  html_document:
    toc: true
    df_print: paged
  word_document:
    toc: true
---

## What this document is

Participants tapped along with a metronome. Partway through some trials, the underlying meter
changed (a "switch") -- either **cued** (an external signal told them when) or **spontaneous** (no
signal; if a switch happens, it's the participant's own doing). We recorded every tap's timing and
loudness and want to know: how does tapping behaviour change around a switch, and can we estimate
*when* a switch happened even when nobody told the participant to switch?

Currently based on one participant (**sub02**) with clean, low-noise recordings; a second
participant's data was excluded (recording was too quiet/noisy to trust).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 6, fig.height = 4, fig.align = "center")

packages <- c("ggplot2", "data.table", "gridExtra")
sapply(packages, require, character.only = TRUE)

theme_set(theme_classic(base_size = 11))
theme_update(text = element_text(size = 11),
             legend.text = element_text(size = 9),
             plot.title = element_text(size = 12, face = "bold"))
```

```{r load-data}
alignment_files <- list.files('../save/derived',
                               pattern = '_click_tap_alignment\\.csv$',
                               recursive = TRUE, full.names = TRUE)
metro <- rbindlist(lapply(alignment_files, function(f) as.data.table(read.csv(f))))

# pandas writes booleans as "True"/"False" strings; read.csv keeps them as
# character, not logical
metro[, missed_beat := as.logical(missed_beat)]
metro[, low_confidence_trial := as.logical(low_confidence_trial)]

# sub01's recording was too quiet/noisy to trust (see project notes) -- excluded
metro <- metro[subject == 'sub06']

# convert the experimenter's cue time (seconds) to a beat count, using each
# trial's own tempo, so trials at different speeds can be compared on the
# same "beats from the switch" axis
metro[, switch_beat := switch_time * bpm / 60]
metro[, click_rel := click_idx - switch_beat]

model_fits <- as.data.table(read.csv('../save/derived/sub06/sub06_spontaneous_tau_fit.csv'))
model_fits <- model_fits[, .(trial, tau = tau_mean, tau_ci_width)]

cued_fits <- as.data.table(read.csv('../save/derived/sub06/sub06_cued_tau_fit.csv'))
cued_fits[, converged := as.logical(converged)]  # pandas writes "True"/"False" strings, not logical

# group-level parameter comparisons (cued vs spontaneous, meter 2 vs 3),
# computed in 02_fit_model.ipynb via a paired Monte Carlo resample of
# posterior draws across trials -- see Part 4 below
group_comparisons <- as.data.table(read.csv('../save/derived/sub06/sub06_group_comparisons.csv'))
group_comparisons[, credible := as.logical(credible)]

posterior_draws <- as.data.table(read.csv('../save/derived/sub06/sub06_posterior_draws.csv'))
posterior_draws[, converged := as.logical(converged)]

# how many trials/participants we'd need to detect various effects, computed
# by Monte Carlo simulation in power_analysis.ipynb -- see Part 5 below
power_trials <- as.data.table(read.csv('results/power_required_trials.csv'))
power_participants <- as.data.table(read.csv('results/power_required_participants.csv'))
```

# Part 1: One trial, up close

Before any pooling or modelling, here's a single cued trial in full: every tap's timing (top) and
loudness (bottom), with the red line marking exactly when the participant was cued to switch meter.

```{r cold-open, fig.height=5}
COLD_OPEN_TRIAL <- 1
one_trial <- metro[condition == 'cued' & trial == COLD_OPEN_TRIAL]
one_switch_beat <- one_trial$switch_beat[1]

p_timing <- ggplot(one_trial[missed_beat == FALSE], aes(x = click_idx, y = asynchrony_s)) +
  geom_line(color = "#4C72B0") +
  geom_point(color = "#4C72B0", size = 1) +
  geom_vline(xintercept = one_switch_beat, color = "firebrick") +
  labs(x = NULL, y = "asynchrony (s)",
       title = paste0("Trial ", COLD_OPEN_TRIAL, ": timing"),
       subtitle = "Red line = when the switch was cued")

p_loud <- ggplot(one_trial[missed_beat == FALSE], aes(x = click_idx, y = rms_db)) +
  geom_line(color = "#55A868") +
  geom_point(color = "#55A868", size = 1) +
  geom_vline(xintercept = one_switch_beat, color = "firebrick") +
  labs(x = "beat number in the trial", y = "tap loudness (dB)",
       title = paste0("Trial ", COLD_OPEN_TRIAL, ": loudness"))

gridExtra::grid.arrange(p_timing, p_loud, ncol = 1)
```

Look at the two panels around the red line. In the **timing** panel, it's hard to tell anything
happened at all -- the tap-to-click offset looks about as noisy before as after. In the
**loudness** panel, though, there's a visible change in *pattern*: before the switch, every other
beat is noticeably louder (a clear accent, alternating strong-weak); after it, that particular
alternation is gone.

## Is this trial typical?

```{r descriptives-table}
desc <- metro[, .(n_trials = length(unique(trial))), by = .(condition, meter)]
setorder(desc, condition, meter)
knitr::kable(desc, col.names = c("Condition", "Meter", "N trials"))
```

20 trials total: 10 **cued** (an instruction tells the participant when to switch) and 10
**spontaneous** (no instruction), evenly split between the two meters tested (2 and 3 beats per
bar). Tempo ranged `r round(min(metro$bpm))`-`r round(max(metro$bpm))` BPM across trials. Pooled
across every trial, timing stays centred on zero with no obvious switch-related structure,
confirming the single trial above wasn't a fluke -- **timing alone doesn't show a reliable switch
signal.**

```{r accent-overview}
amp <- metro[missed_beat == FALSE]
amp[, bar_position := click_idx %% meter]

agg_amp <- amp[, .(mean_db = mean(rms_db, na.rm = TRUE),
                    se = sd(rms_db, na.rm = TRUE) / sqrt(.N)),
               by = .(condition, meter, bar_position)]

ggplot(agg_amp, aes(x = factor(bar_position), y = mean_db, fill = condition)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6) +
  geom_errorbar(aes(ymin = mean_db - se, ymax = mean_db + se),
                position = position_dodge(width = 0.7), width = 0.2) +
  facet_wrap(~ meter, scales = "free_x", labeller = label_both) +
  labs(x = "position within the bar", y = "mean tap loudness (dB)", fill = NULL,
       title = "Tap loudness by beat position",
       subtitle = "Pooled across all trials")
```

Loudness, on the other hand, generalises: across trials and both meters there's a consistent
loudness contrast between beat positions (a louder "downbeat"). **This is why the model below is
built entirely on loudness, not timing** -- it's the channel that actually carries the switch
signal.

# Part 2: What the model does

We fit a statistical model to each trial's tap-loudness pattern. In plain terms, the model looks for
**the point in the trial where the participant's tapping pattern changes from one meter's accent to
the other's**, and reports:

- **`tau`** -- **when** the change happened, given as a beat number in the trial (or, for cued
  trials, we can compare it directly to when the instruction was given).
- **how confident** the model is in that estimate, from the *width* of `tau`'s credible interval --
  a trial with a clear, well-defined switch gives a narrow, tight estimate; a trial where the data
  don't show a clear switch (or don't show one clearly enough to pin down) gives a wide, uncertain
  estimate instead of a falsely precise one.

Concretely, each beat's loudness (in dB) is converted to a log-amplitude number, and that's the one
value the model fits at every beat -- expected to sit near one meter's accent pattern before the
switch and the other's after it, blending smoothly in between. (Full formulas in
`../docs/model_math.md`.)

The model allows the switch to be gradual, but doesn't assume it has to be -- checking this
directly against the data (comparing a smooth-transition version of the model against a version
where the switch is instantaneous) shows real switches are close to instantaneous: tapping style
changes over roughly a beat, not gradually across many beats. Either way, the model only uses the
loudness/accent information described in Part 1 -- not raw timing -- to find the change point.

```{r model-schematic, fig.width=8, fig.height=5}
# Illustrative only -- a smooth, textbook-shaped example, not a fit to any
# real trial -- just to make "switch, weighted blend" concrete. Width is set
# small/sharp here to match what fitting to real data actually shows (see
# text above).
n_demo <- 0:50
tau_demo <- 25
width_demo <- 0.5  # how many beats the transition takes to unfold
w_demo <- 1 / (1 + exp(-(n_demo - tau_demo) / width_demo))

# toy 2-beat and 3-beat accent templates (arbitrary, just for illustration)
g2 <- c(1, -1)
g3 <- c(1, -1, -0.9)
loudness_demo <- (1 - w_demo) * g2[(n_demo %% 2) + 1] + w_demo * g3[(n_demo %% 3) + 1]

df_schematic <- data.table(n = n_demo, w = w_demo, loudness = loudness_demo)

p1 <- ggplot(df_schematic, aes(x = n, y = w)) +
  geom_line(color = "#4C72B0", linewidth = 1) +
  geom_vline(xintercept = tau_demo, color = "firebrick", linetype = "dashed") +
  annotate("text", x = tau_demo + 1, y = 0.15, label = "tau", color = "firebrick",
           hjust = 0, fontface = "italic") +
  labs(x = "beat number", y = "switch progress\n(0 = old meter, 1 = new meter)",
       title = "How much has the switch happened by this beat?")

p2 <- ggplot(df_schematic, aes(x = n, y = loudness)) +
  geom_line(color = "#55A868", linewidth = 1) +
  geom_vline(xintercept = tau_demo, color = "firebrick", linetype = "dashed") +
  labs(x = "beat number", y = "expected tap loudness",
       title = "Resulting accent pattern (what the model fits to the data)",
       subtitle = "Old meter's accent fades out, new meter's accent fades in, centred on tau")

gridExtra::grid.arrange(p1, p2, ncol = 1)
```

*(Illustrative example, not a fit to real data.)* Top: the model's "switch progress" curve -- a
smooth S-shape that goes from 0 (still in the old meter) to 1 (fully in the new meter), centred on
`tau`. Bottom: this progress curve blends the old meter's accent pattern into the new one's, which
is what the model actually compares against the real tap-loudness data. The model can fit either a
narrow (quick) or wide (gradual) S-curve, whichever the data support -- and as noted above, real
trials consistently favour a narrow one, close to an instant switch.

We do **not** detail the model's internal mechanics beyond this here; see `../docs/model_math.md`
and `02_fit_model.ipynb` for the full specification.

**Validation.** Before trusting the model on the harder spontaneous case, we checked it against the
cued trials, where we know when the *instruction* to switch was given. `tau` isn't meant to match
that instruction time exactly -- it's fit purely from tap loudness, so it measures when the
participant's *tapping itself* actually changed, which should come a little after the instruction
(reaction time, plus a moment to settle into the new pattern). That's exactly what we see below:
across 9 of the 10 cued trials (one trial's fit didn't converge reliably and is excluded -- flagged
in the table), `tau` tracks the instruction time closely with a consistent positive lag, never
negative.

```{r cued-validation-table}
cued_fits[, lag_beats := tau_mean - switch_beat]
tbl_cued <- cued_fits[order(trial), .(trial, meter,
                                       "instruction time (beats)" = round(switch_beat, 1),
                                       "model's tau (beats)" = round(tau_mean, 1),
                                       "lag (tau - instruction)" = round(lag_beats, 1),
                                       "converged?" = ifelse(converged, "yes", "NO -- excluded below"))]
knitr::kable(tbl_cued)
```

```{r cued-validation-plot, fig.width=6, fig.height=5}
cued_ok <- cued_fits[converged == TRUE]
r <- round(cor(cued_ok$tau_mean, cued_ok$switch_beat), 2)
lag_mean <- round(mean(cued_ok$lag_beats), 2)

ggplot(cued_fits, aes(x = switch_beat, y = tau_mean, color = converged)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
  geom_errorbar(aes(ymin = tau_lo, ymax = tau_hi), width = 0.5, alpha = 0.6) +
  geom_point(size = 2.5) +
  scale_color_manual(values = c("TRUE" = "#4C72B0", "FALSE" = "firebrick"),
                      labels = c("TRUE" = "converged", "FALSE" = "not converged (excluded)")) +
  labs(x = "instruction time (beats)", y = "model's tau (beats)", color = NULL,
       title = "Model's tau vs. the actual cue time, per cued trial",
       subtitle = paste0("Converged trials: correlation = ", r, ", mean lag = ", lag_mean, " beats\n",
                          "Dashed line = perfect agreement (tau == instruction time)"))
```

# Part 3: Spontaneous switches

With no cue, there's no obvious external marker to tell us when (or whether) a participant
switched on their own. This is exactly where the model's `tau` estimate helps: for each trial, we
can ask *when the model thinks the participant changed*, and *how sure it is* (from the width of
`tau`'s credible interval).

```{r spont-confidence-table}
CI_WIDTH_THRESHOLD <- 15  # beats; below this, treat tau as well-identified
model_fits[, confident_switch := tau_ci_width < CI_WIDTH_THRESHOLD]
tbl <- model_fits[, .(trial, "estimated switch (beat)" = round(tau, 1),
                       "94% CI width (beats)" = round(tau_ci_width, 1),
                       "model confidence" = ifelse(confident_switch, "confident switch", "unclear / no switch"))]
knitr::kable(tbl)
```

`r sum(model_fits$confident_switch)` of the 10 spontaneous trials show a confident, well-identified
switch; the rest are ambiguous (a wide `tau` credible interval means the data don't clearly pin down
when -- or whether -- the participant's tapping pattern actually changed). We only use the confident
trials below -- mixing in the ambiguous ones would add noise from switch estimates the model itself
is unsure about.

## A model-free check: does the accent pattern itself show a switch?

Before trusting the model's `tau` estimates, it's worth asking whether a switch is visible directly
in the raw loudness data, with no fitting at all. For each trial, we slide an 8-beat window along the
trial and compute one simple number: **how much loudness varies across the different positions in
the bar, right now** (the gap between the loudest and quietest position, in dB, using only the taps
in that window). If a participant has a clear accent pattern, this number is high; if it fades to a
flat, unaccented tapping style, this number drops toward zero. A clear, sustained drop is a
model-free signature of a switch.

```{r model-free-contrast-fn}
# rolling "how much does loudness vary by position right now" statistic,
# no model fitting involved -- just grouping and a max-min gap in a
# sliding window of raw data
rolling_accent_contrast <- function(dt, window_beats = 8) {
  beats <- seq(min(dt$click_idx), max(dt$click_idx))
  rbindlist(lapply(beats, function(ctr) {
    w <- dt[click_idx >= ctr - window_beats & click_idx < ctr + window_beats]
    by_pos <- w[, .(mean_db = mean(rms_db, na.rm = TRUE)), by = bar_position]
    if (nrow(by_pos) < uniqueN(dt$bar_position)) return(NULL)
    data.table(click_idx = ctr, contrast_db = max(by_pos$mean_db) - min(by_pos$mean_db))
  }))
}

spont_amp_trials <- metro[condition == 'spontaneous' & missed_beat == FALSE]
spont_amp_trials[, bar_position := click_idx %% meter]

contrast_by_trial <- rbindlist(lapply(unique(spont_amp_trials$trial), function(tr) {
  out <- rolling_accent_contrast(spont_amp_trials[trial == tr])
  out[, trial := tr]
  out
}))

contrast_by_trial <- merge(contrast_by_trial, model_fits[, .(trial, tau, confident_switch)], by = 'trial')
```

```{r model-free-contrast-plot, fig.height=6}
ggplot(contrast_by_trial) +
  geom_line(aes(x = click_idx, y = contrast_db), color = "#8172B2") +
  geom_vline(aes(xintercept = tau), color = "firebrick", linetype = "dashed") +
  facet_grid(trial ~ ., scales = "free_y") +
  labs(x = "beat number in the trial", y = "accent contrast (dB)",
       title = "Model-free accent contrast over each trial",
       subtitle = "Dashed line = model's estimated switch beat, for comparison")
```

Where the model reports a confident switch, the accent contrast typically starts high and shows a
clear, sustained drop close to the model's estimated `tau` -- a useful independent check that the
model isn't just fitting noise. Note that individual taps still look "zig-zaggy" by eye even after a
real collapse (beat-to-beat noise is large relative to the shrunk accent), so a raw per-tap plot
alone can look similar before and after a switch -- the rolling contrast statistic is what actually
reveals the change, by averaging out that noise.

Where the model was unsure, there are two different reasons, visible directly in this plot:
trials where the contrast **never gets very high to begin with** (e.g. trial 13, which starts under
2 dB) have no strong initial accent to switch away from, so there's nothing for the model to detect
a collapse in; trials where the contrast drops only partway and not all the way to zero (e.g. trial
18) show a weaker, less complete change. Both are genuine features of the data, not a shortcoming of
the fitting procedure -- they're exactly why the model correctly reports low confidence there rather
than forcing a precise answer.

## Aligning spontaneous trials on the model's estimated switch

```{r spont-aligned-plot, fig.height=5}
spont_fitted <- metro[condition == 'spontaneous']
spont_fitted <- merge(spont_fitted, model_fits[, .(trial, tau, confident_switch)], by = 'trial')
spont_fitted[, click_rel := click_idx - tau]

spont_confident <- spont_fitted[confident_switch == TRUE]

ggplot(spont_confident) +
  geom_line(aes(x = click_rel, y = asynchrony_s), color = "#55A868") +
  geom_vline(xintercept = 0, color = "firebrick") +
  facet_grid(trial ~ .) +
  labs(x = "beats relative to the model's estimated switch point", y = "asynchrony (s)",
       title = "Confident spontaneous switches, aligned",
       subtitle = "Red line = model's estimated switch beat, per trial")
```

The main takeaway here is methodological: the model can pick out *which* spontaneous trials had a
clear switch and roughly *when*, purely from the loudness/accent pattern, without ever being told a
switch happened.

## A closer look: a small timing bump right after the switch

Timing wasn't reliable enough to detect a switch on its own (Part 1), but now that we have `tau`
from the loudness model, it's worth looking at timing *around* that known point more closely than
before. Using a narrow sliding window (much finer than earlier passes), there's a small, localised
bump in timing error right after the switch: tap-to-click error dips exactly at the switch, then
rises again for a few beats, before settling back down.

```{r timing-bump-fn}
# Finer-grained than the earlier pooled view: a narrow window (in beats) so a
# short-lived bump isn't averaged away by a wider one.
sliding_window_narrow <- function(dt, half_width_beats = 2, step_beats = 1) {
  centers <- seq(min(dt$click_rel), max(dt$click_rel), by = step_beats)
  rbindlist(lapply(centers, function(ctr) {
    in_win <- dt[click_rel >= ctr - half_width_beats & click_rel < ctr + half_width_beats]
    per_trial <- in_win[, .(trial_stat = mean(abs(asynchrony_s), na.rm = TRUE),
                             n = sum(!is.na(asynchrony_s))), by = .(subject, trial)]
    per_trial <- per_trial[n >= 3]
    if (nrow(per_trial) == 0) return(NULL)
    data.table(click_rel = ctr,
               mean_stat = mean(per_trial$trial_stat, na.rm = TRUE),
               se = sd(per_trial$trial_stat, na.rm = TRUE) / sqrt(nrow(per_trial)),
               n_trials = nrow(per_trial))
  }))
}

by_rel_bump <- sliding_window_narrow(metro[condition == 'cued' & click_rel > -15 & click_rel < 15])
```

```{r timing-bump-plot}
ggplot(by_rel_bump, aes(x = click_rel, y = mean_stat)) +
  geom_ribbon(aes(ymin = mean_stat - se, ymax = mean_stat + se), alpha = 0.2, fill = "#C44E52") +
  geom_line(color = "#C44E52") +
  geom_vline(xintercept = 0, color = "firebrick", linetype = "dashed") +
  labs(x = "beats relative to the cued switch", y = "average timing error (s)",
       title = "A local timing bump shortly after the switch",
       subtitle = "Narrow window (cued trials); shaded band = standard error across trials")
```

This is a small effect on one participant's data, so it isn't something to draw conclusions from
yet -- but it's worth keeping an eye on once more participants are added: if it holds up, it would
suggest a brief timing wobble a beat or two after the switch itself (distinct from the broader
pre-vs-post trend), which could be a genuine signature of the participant re-locking onto the new
meter right after adopting its accent.

# Part 4: Comparing conditions and meters

Beyond `tau` itself, the model fits three other parameters per trial: `s` (how sharp the switch
is), `sigma_sw` (extra noise right at the switch), and `sigma_0` (baseline loudness noise,
unrelated to the switch). Do any of these differ between cued and spontaneous trials, or between
meter 2 and meter 3? Comparing posterior *means* alone can be misleading -- it ignores how
uncertain each trial's own estimate is. Instead, `02_fit_model.ipynb` computes a proper 94%
credible interval on each group difference (by resampling posterior draws across trials within
each group many times, building up a distribution of "the group's mean" that respects each
trial's own uncertainty, then differencing). A difference only counts as credible if that interval
excludes zero.

```{r group-comparison-table}
tbl_group <- copy(group_comparisons)
tbl_group[, group_a_mean := fifelse(comparison == "cued - spontaneous", cued_mean, X2_mean)]
tbl_group[, group_b_mean := fifelse(comparison == "cued - spontaneous", spontaneous_mean, X3_mean)]
tbl_group[, group_a_label := fifelse(comparison == "cued - spontaneous", "cued", "meter 2")]
tbl_group[, group_b_label := fifelse(comparison == "cued - spontaneous", "spontaneous", "meter 3")]

tbl_display <- tbl_group[, .(
  Comparison = paste(group_a_label, "vs", group_b_label),
  Parameter = parameter,
  "Group A mean" = round(group_a_mean, 3),
  "Group B mean" = round(group_b_mean, 3),
  Difference = round(diff_mean, 3),
  "94% credible interval" = paste0("[", round(hdi_94_lo, 3), ", ", round(hdi_94_hi, 3), "]"),
  "Credible?" = ifelse(credible, "yes", "no")
)]
knitr::kable(tbl_display)
```

Two things stand out. `tau` differs credibly between cued and spontaneous -- but that's not a new
result, it's the same reaction-lag effect already reported in Part 2 (cued trials' `tau` sits
`switch_beat` + ~2.3 beats later; spontaneous trials have no `switch_beat` to add that to). `s` and
`sigma_sw` show no credible difference anywhere: switches are equally sharp and equally noisy at
the switch, regardless of condition or meter, as far as this data can tell.

The one genuinely new result is `sigma_0`: **meter-2 trials are noisier at baseline than meter-3
trials**, regardless of condition -- consistent across nearly every individual trial, not driven
by one outlier.

```{r sigma0-by-meter-plot}
sigma0_per_trial <- posterior_draws[converged == TRUE,
  .(mean_sigma0 = mean(sigma_0)), by = .(trial, meter, condition)]

diff_row <- group_comparisons[comparison == "2 - 3" & parameter == "sigma_0"]

ggplot(sigma0_per_trial, aes(x = factor(meter), y = mean_sigma0)) +
  geom_boxplot(width = 0.4, outlier.shape = NA, fill = "grey90") +
  geom_jitter(aes(color = condition), width = 0.08, size = 2.5) +
  labs(x = "meter", y = "sigma_0 (baseline noise, log-amplitude units)", color = NULL,
       title = "Baseline tapping noise is higher in meter-2 trials",
       subtitle = paste0("meter 2 - meter 3 = ", round(diff_row$diff_mean, 3),
                          ", 94% credible interval [", round(diff_row$hdi_94_lo, 3), ", ",
                          round(diff_row$hdi_94_hi, 3), "]"))
```

This is unrelated to the switch itself -- `sigma_0` describes general tap-to-tap loudness
consistency, not anything specific to *when* the meter changes. A plausible reading: a 2-beat bar
gives the tapper less internal structure to lock onto (just strong-weak) than a 3-beat bar
(strong-weak-weak), making steady, consistent accenting harder to sustain -- but with one
participant this is a hypothesis worth checking with more data, not a conclusion.

# Part 5: How much data would we actually need?

Part 4's comparisons came back mostly non-credible with the ~5 trials/cell we have now (meter 2/3
x cued/spontaneous). That's ambiguous -- "no real effect" or just "not enough data"? `power_analysis.ipynb`
simulates new datasets at the same trial-to-trial noise seen in sub02's fits to answer that; full
method there, headline numbers below.

```{r power-trials-table}
tbl_power_trials <- power_trials[, .(
  Parameter = parameter,
  Effect = effect,
  "Effect size" = paste0(vs_observed, " the observed diff"),
  "Standardized effect (Cohen's d)" = round(cohens_d, 2),
  "Trials/cell needed for 80% power" = n_per_cell_for_80pct_power
)]
knitr::kable(tbl_power_trials)
```

Most effects *actually observed* so far need well more than the current ~5 trials/cell to reach
80% power (e.g. 104 for `tau`'s meter effect), dropping fast for a bigger true effect (12 at 3x) --
so most Part 4 non-findings are inconclusive, not evidence of no effect. `sigma_0`'s meter effect
needed no such adjustment (already ~80% powered at n=5), which is a useful sanity check on the
method. One new result: the **meter x condition interaction on `tau`** (d = 0.83) is bigger than
either main effect and was never tested before -- worth a dedicated follow-up (~48 trials/cell to
confirm).

**Suggestion:** **~20 trials/cell (80/participant)** as a practical default, enough for a
moderate-to-large effect (2-3x observed) on most parameters above; size specifically to the `tau`
interaction row if that's the actual target.

Separately: if switch "selectivity" (`-log(s)`) scales with musical training, how many
*participants* would that take? (Needs an assumption about between-participant variance, which one
participant can't measure -- see the notebook.)

```{r power-participants-table}
tbl_power_participants <- power_participants[, .(
  "True correlation (r)" = true_r,
  "Trials/participant" = trials_per_participant,
  "Participants needed for 80% power" = n_participants_for_80pct_power
)]
knitr::kable(tbl_power_participants)
```

Participant count is the real lever here, not trials/participant (which barely moves the
requirement): a moderate effect (r = 0.5) needs roughly **30 participants**, a strong one (r = 0.7)
**~15** -- keep sessions around the current length and prioritise recruiting more participants.

# Summary

- Tapping timing looks noisy and undifferentiated around a switch, both in a single trial and
  pooled across all of them -- **loudness, not timing, is where the switch signal actually lives**,
  which is why the model is built entirely on tap loudness.
- The switch-detection model was validated against cued trials (known instruction time) before being
  trusted on spontaneous trials: `tau` tracks the instruction time closely (correlation 0.94) with a
  small, consistent, always-positive lag (~2.3 beats on average) -- a sensible reaction/adoption
  delay, not the cue time itself, which is what the model is actually built to measure. Real
  switches also turned out to be close to instantaneous rather than gradual, once checked directly.
- For spontaneous trials, the model both estimates *when* a switch happened and *how confident* it
  is in that estimate -- `r sum(model_fits$confident_switch)` in 10 spontaneous trials show a
  confident, well-identified switch.
- A model-free check (raw accent contrast, no fitting) confirms the pattern independently: confident
  trials show a real contrast collapse near the model's estimated switch; unclear trials show either
  a weak starting accent or an incomplete collapse -- genuine data features, not fitting artifacts.
- All results so far are based on a single participant; more data is needed before drawing general
  conclusions.
- A power analysis of the full 2x2 (meter x condition) design suggests roughly **20 trials/cell
  (80/participant)** to properly resolve most `tau`/`s` effects at 2-3x their currently-observed
  size, and hints at a **`tau` meter x condition interaction** worth a dedicated follow-up. Across
  participants, roughly **30 participants** (session length largely unchanged) would be needed to
  detect a moderate musical-training effect on switch selectivity.
